---
title: "Reproducible Research: Peer Assessment 1"
output: 
  html_document:
    keep_md: true
---


## Loading and preprocessing the data

```{r,message=FALSE}
library(dplyr)
activity<-tbl_df(read.csv(file = "activity.csv"))
```

## What is mean total number of steps taken per day?

Total number of steps taken per day

```{r,warning=FALSE,fig.width=10}

steps_by_date<-summarise(group_by(activity,date), sum_steps=sum(steps))
print(steps_by_date,n=10e5)


library(ggplot2)

ggplot(steps_by_date, aes(date,sum_steps)) +
       geom_histogram(stat="identity",fill="lightblue", colour="black",width=1) +
       theme(axis.text.x = element_text(angle=60, hjust=1, vjust=1,size=rel(0.8)))+
       ylab ("Number of steps") +
       xlab ("Date")+
       ggtitle("Total number of steps taken each day")
```

Mean and median of the total number of steps taken per day
```{r,}
options(digits = 2)
steps_by_date<-summarise(group_by(activity,date),
                         mean_steps= mean(steps,na.rm = TRUE),
                         median_steps=median(steps,na.rm = TRUE))

print(steps_by_date,n=10e5)
```

## What is the average daily activity pattern?
```{r,fig.width=10}
steps_by_interval<-summarise(group_by(activity,interval), 
                             mean_steps= mean(steps,na.rm = TRUE))

ggplot(steps_by_interval, aes(interval,mean_steps)) +
       geom_line(colour="red",size=1) +
       ylab ("Number of steps") +
       xlab ("Interval")+
       ggtitle("Average of number of steps each interval")

max_steps_by_interval<-summarise(group_by(activity,interval), 
                            max_steps= max(steps,na.rm = TRUE))
```
Max of number of steps each interval
```{r,fig.width=10}
ggplot(max_steps_by_interval, aes(interval,max_steps)) +
       geom_line(colour="red",size=1) +
       ylab ("Number of steps") +
       xlab ("Interval")+
       ggtitle("Max of number of steps each interval")
```

## Imputing missing values

Total number of missing values in the dataset
```{r,}
sum(!complete.cases(activity))
```

Filling  missing values in the dataset with the mean of 5-minute interval
```{r,}
activitynew <- activity  ## New dataset
for (i in 1:length(activitynew$steps)) {
          if (is.na(activitynew[i,"steps"]))
                 ## Assigned the  mean of 5-minute interval if steps = NA at new dataset 
                 activitynew[i,"steps"] <- select(filter(steps_by_interval, interval==as.integer(activitynew[i,"interval"])),mean_steps)    
                 
}

```
Total number of steps taken per day. Filling  missing values
```{r,warning=FALSE,fig.width=10}

steps_by_date<-summarise(group_by(activitynew,date), sum_steps=sum(steps))


library(ggplot2)

ggplot(steps_by_date, aes(date,sum_steps)) +
       geom_histogram(stat="identity",fill="lightblue", colour="black",width=1) +
       theme(axis.text.x = element_text(angle=60, hjust=1, vjust=1,size=rel(0.8)))+
       ylab ("Number of steps") +
       xlab ("Date")+
       ggtitle("Total number of steps taken each day")
```
Mean and median of the total number of steps taken per day. Filling  missing values
```{r,}
options(digits = 2)
steps_by_date<-summarise(group_by(activitynew,date),
                         mean_steps= mean(steps,na.rm = TRUE),
                         median_steps=median(steps,na.rm = TRUE))

print(steps_by_date,n=10e5)
```

## Are there differences in activity patterns between weekdays and weekends?

```{r,warning=FALSE,fig.width=10,fig.height=8}

library(chron)
activitynew$week <- factor(is.weekend(activitynew$date),labels = c("Weekday","Weekend"))

steps_by_interval_w<-summarise(group_by(activitynew,interval,week), 
                             mean_steps= mean(steps,na.rm = TRUE))

ggplot(steps_by_interval_w, aes(interval,mean_steps,week)) +
       geom_line(colour="red",size=1) +
       ylab ("Number of steps") +
       xlab ("Interval")+
       facet_wrap( ~ week,ncol=1)+
       ggtitle("Average of number of steps each interval Weekday / Weekend")

```

